Analysis of Bauer et al. (unpublished) Field experiment: <br> Persistence

### 2022-12-08 R/model_fcs_check.Rmd
Update markdown

---
title: "Analysis of Bauer et al. (unpublished) Field experiment: <br> Favourable Conservation Status (FCS)"
author: "<b>Markus Bauer</b> <br>"
date: "<b>`r format(Sys.time(), '%Y-%m-%d')`</b>"
output:
  github_document:
    toc: true
    toc_depth: 3
    dev: png
    fig_width: 7
    fig_height: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
  )
```
<br/>
<br/>
<b>Markus Bauer</b>

@@

[markus1.bauer@tum.de](mailto:markus1.bauer@tum.de)

@@

```{r load-models, collapse = TRUE}
load(file = here("outputs", "models", "model_fcs_2.Rdata"))
load(file = here("outputs", "models", "model_fcs_full.Rdata"))
load(file = here("outputs", "models", "model_fcs_2_prior.Rdata"))
# BARG 5.A/B/C
#load(file = here("outputs", "models", "model_fcs_2_flat.Rdata"))
m_1 <- m2
m_2 <- m_full
m_prior <- m2_prior
```

@@

```{r analysis-preparation}
posterior1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth15",
      "b_target_typedry_grassland",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
  )
posterior2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth15",
      "b_target_typedry_grassland",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
  )
posterior_prior <- m_prior %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "Intercept",
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth15",
      "b_target_typedry_grassland",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021"
    )
  )
hmc_diagnostics1 <- brms::nuts_params(m_1)
hmc_diagnostics2 <- brms::nuts_params(m_2)
y <- sites$n
yrep1 <- brms::posterior_predict(m_1, draws = 500)
yrep2 <- brms::posterior_predict(m_2, draws = 500)
loo1 <- brms::loo(m_1, save_psis = TRUE, moment_match = FALSE)
loo2 <- brms::loo(m_2, save_psis = TRUE, moment_match = FALSE)
draws1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
draws2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
```

@@

#### Prior predictive check (BARG 1.E)

```{r prior-predictive-check}
bayesplot::mcmc_intervals(
  posterior_prior,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
  ) +
  theme_classic()
```

@@

```{r rhat, collapse = TRUE}
mcmc_rhat(draws1$rhat)
mcmc_rhat(draws2$rhat)
```

@@

```{r ess, collapse = TRUE}
mcmc_neff(neff_ratio(m_1))
mcmc_neff(neff_ratio(m_2))
```

@@

Chosen model:

```{r posteriors-chosen-model}
mcmc_intervals(
  posterior1,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
  ) +
  theme_classic()
```

Second model:

```{r posterors-2nd-model}
mcmc_intervals(
  posterior2,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
  ) +
  theme_classic()
```

>>>>

---
title: "Analysis of Bauer et al. (unpublished) Field experiment: <br> Favourable Conservation Status (FCS)"
author: "<b>Markus Bauer</b> <br>"
date: "<b>`r format(Sys.time(), '%Y-%m-%d')`</b>"
output:
  github_document:
    toc: true
    toc_depth: 3
    dev: png
    fig_width: 7
    fig_height: 5
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
  )
```

<br/> <br/> <b>Markus Bauer</b>

@@

[markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de)

@@

```{r load-models, collapse = TRUE}
load(file = here("outputs", "models", "model_fcs_2.Rdata"))
load(file = here("outputs", "models", "model_fcs_full.Rdata"))
load(file = here("outputs", "models", "model_fcs_2_prior.Rdata"))
# BARG 5.A/B/C
load(file = here("outputs", "models", "model_fcs_2_flat.Rdata"))
m_1 <- m2
m_2 <- m_full
m_prior <- m2_prior
m_flat <- m2_flat
```

@@

```{r analysis-preparation}
# Chose variables
variables <- c(
      "Intercept",
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth15",
      "b_target_typedry_grassland",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
# Subset draws
posterior1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
posterior2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
posterior_prior <- m_prior %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
posterior_flat <- m_flat %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
# R hat
rhat1 <- rhat(m_1)
rhat2 <- rhat(m_2)
# NEFF ratio
neff1 <- neff_ratio(m_1)
neff2 <- neff_ratio(m_2)
# Long format of draws
hmc_diagnostics1 <- brms::nuts_params(m_1)
hmc_diagnostics2 <- brms::nuts_params(m_2)
y <- sites$n
# Posterior predictive distribution
yrep1 <- brms::posterior_predict(m_1, draws = 500)
yrep2 <- brms::posterior_predict(m_2, draws = 500)
# Leave-one-out cross validation based on posterior likelihood
loo1 <- brms::loo(m_1, save_psis = TRUE, moment_match = FALSE)
loo2 <- brms::loo(m_2, save_psis = TRUE, moment_match = FALSE)
# Summary statistics
draws1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
draws2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
```

@@

@@

```{r rhat, collapse = TRUE}
mcmc_rhat(rhat1)
mcmc_rhat(rhat2)
```

@@

```{r ess, collapse = TRUE}
mcmc_neff(neff1)
mcmc_neff(neff2)
```

@@

```{r posteriors, fig.height = 7}
combined <- rbind(
  bayesplot::mcmc_intervals_data(posterior1, prob = 0.66, prob_outer = 0.95),
  bayesplot::mcmc_intervals_data(posterior2, prob = 0.66, prob_outer = 0.95),
  bayesplot::mcmc_intervals_data(posterior_prior, prob = 0.66, prob_outer = 0.95),
  bayesplot::mcmc_intervals_data(posterior_flat, prob = 0.66, prob_outer = 0.95)
  )
combined$model <- rep(
  c("m_1", "m_2", "m_flat", "m_prior"),
  each = posterior1$"1" %>% as_tibble() %>% ncol()
  )
pos <- position_nudge(
  y = if_else(
    combined$model == "m_2", -.2, if_else(
      combined$model == "m_prior", -.4, if_else(
        combined$model == "m_flat", -.6, 0
        )
      )
    )
  )
ggplot(data = combined, aes(x = m, y = forcats::fct_rev(factor(parameter)), color = model)) + 
  geom_vline(xintercept = 0, color = "grey") +
  geom_linerange(aes(xmin = l, xmax = h), position = pos, linewidth = 2) +
  geom_linerange(aes(xmin = ll, xmax = hh), position = pos) +
  geom_point(position = pos, color = "black") +
  coord_cartesian(xlim = c(-1, .6)) +
  bayesplot::theme_default() +
  ggtitle("Posterior distbributions (mean, CI66, CI95)")
```

### 2022-12-08 R/model_persistence_check.Rmd
Update markdown

```{r load-models, collapse = TRUE, message = FALSE}
load(file = here("outputs", "models", "model_persistence_2.Rdata"))
load(file = here("outputs", "models", "model_persistence_full.Rdata"))
load(file = here("outputs", "models", "model_persistence_2_prior.Rdata"))
# BARG 5.A/B/C
#load(file = here("outputs", "models", "model_persistence_2_flat.Rdata"))
m_1 <- m2
m_2 <- m_full
m_prior <- m2_prior
```

@@

```{r rhat, collapse = TRUE}
mcmc_rhat(draws1$rhat)
mcmc_rhat(draws2$rhat)
```

@@

Chosen model:

```{r posteriors-chosen-model}
mcmc_intervals(
  posterior1,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
  ) +
  theme_classic()
```

Second model:

```{r posterors-2nd-model}
bayesplot::mcmc_intervals(
  posterior2,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
  ) +
  theme_classic()
```

>>>>

```{r load-models, collapse = TRUE, message = FALSE}
load(file = here("outputs", "models", "model_persistence_2.Rdata"))
load(file = here("outputs", "models", "model_persistence_full.Rdata"))
load(file = here("outputs", "models", "model_persistence_2_prior.Rdata"))
# BARG 5.A/B/C
load(file = here("outputs", "models", "model_persistence_2_flat.Rdata"))
m_1 <- m2
m_2 <- m_full
m_prior <- m2_prior
m_flat <- m2_flat
```

@@

```{r rhat, collapse = TRUE}
mcmc_rhat(rhat1)
mcmc_rhat(rhat1)
```

@@

```{r posteriors, fig.height = 7}
combined <- rbind(
  bayesplot::mcmc_intervals_data(posterior1, prob = 0.66, prob_outer = 0.95),
  bayesplot::mcmc_intervals_data(posterior2, prob = 0.66, prob_outer = 0.95),
  bayesplot::mcmc_intervals_data(posterior_prior, prob = 0.66, prob_outer = 0.95),
  bayesplot::mcmc_intervals_data(posterior_flat, prob = 0.66, prob_outer = 0.95)
  )
combined$model <- rep(
  c("m_1", "m_2", "m_flat", "m_prior"),
  each = posterior1$"1" %>% as_tibble() %>% ncol()
  )
pos <- position_nudge(
  y = if_else(
    combined$model == "m_2", -.2, if_else(
      combined$model == "m_prior", -.4, if_else(
        combined$model == "m_flat", -.6, 0
        )
      )
    )
  )
ggplot(data = combined, aes(x = m, y = forcats::fct_rev(factor(parameter)), color = model)) + 
  geom_vline(xintercept = 0, color = "grey") +
  geom_linerange(aes(xmin = l, xmax = h), position = pos, linewidth = 2) +
  geom_linerange(aes(xmin = ll, xmax = hh), position = pos) +
  geom_point(position = pos, color = "black") +
  coord_cartesian(xlim = c(-1, 1)) +
  bayesplot::theme_default() +
  ggtitle("Posterior distbributions (mean, CI66, CI95)")
```

### 2022-12-08 R/model_recovery_check.Rmd
Update markdown

```{r load-models, collapse = TRUE, message = FALSE}
load(file = here("outputs", "models", "model_recovery_2.Rdata"))
load(file = here("outputs", "models", "model_recovery_full.Rdata"))
load(file = here("outputs", "models", "model_recovery_2_prior.Rdata"))
# BARG 5.A/B/C
load(file = here("outputs", "models", "model_recovery_2_flat.Rdata"))
m_1 <- m2
m_2 <- m2_flat
#m_2 <- m_full
m_prior <- m2_prior
```

@@

```{r histograms}
#p1 <- ppc_stat(y, yrep1, binwidth = 0.001)
#p2 <- ppc_stat(y, yrep2, binwidth = 0.001)
#p1 / p2
#ppc_stat_grouped(y, yrep1, group = sites$site, binwidth = 0.001)
#ppc_stat_grouped(y, yrep2, group = sites$site, binwidth = 0.001)
#p1 <- ppc_stat_grouped(y, yrep1, group = sites$exposition, binwidth = 0.001)
#p2 <- ppc_stat_grouped(y, yrep2, group = sites$exposition, binwidth = 0.001)
#p1 / p2
#ppc_stat_grouped(y, yrep1, group = sites$survey_year_fct, binwidth = 0.001)
#ppc_stat_grouped(y, yrep2, group = sites$survey_year_fct, binwidth = 0.001)
#p1 <- ppc_stat_grouped(y, yrep1, group = sites$target_type, binwidth = 0.001)
#p2 <- ppc_stat_grouped(y, yrep2, group = sites$target_type, binwidth = 0.001)
#p1 / p2
#p1 <- ppc_stat_grouped(y, yrep1, group = sites$seed_density, binwidth = 0.001)
#p2 <- ppc_stat_grouped(y, yrep2, group = sites$seed_density, binwidth = 0.001)
#p1 / p2
#p1 <- ppc_stat_grouped(y, yrep1, group = sites$sand_ratio, binwidth = 0.001)
#p2 <- ppc_stat_grouped(y, yrep2, group = sites$sand_ratio, binwidth = 0.001)
#p1 / p2
#p1 <- ppc_stat_grouped(y, yrep1, group = sites$substrate_depth, binwidth = 0.001)
#p2 <- ppc_stat_grouped(y, yrep2, group = sites$substrate_depth, binwidth = 0.001)
#p1 / p2
```

@@

```{r loo}
loo1
loo2
#plot(loo1)
#plot(loo2)
```

@@

```{r loo-pit}
#p1 <- bayesplot::ppc_loo_pit_overlay(y, yrep1, lw = weights(loo1$psis_object))
#p2 <- bayesplot::ppc_loo_pit_overlay(y, yrep2, lw = weights(loo2$psis_object))
#p1 / p2

@@

Chosen model:

```{r posteriors}
combined <- rbind(
  mcmc_intervals_data(posterior1, prob = 0.66, prob_outer = 0.95),
  mcmc_intervals_data(posterior2, prob = 0.66, prob_outer = 0.95)
  )
combined$model <- rep(
  c("Model 1", "Model 2"),
  each = ncol(posterior1)
  )
pos <- position_nudge(y = ifelse(combined$model == "Model 2", 0, 0.1))
ggplot(data = combined, aes(x = m, y = parameter, color = model)) + 
  geom_linerange(aes(xmin = l, xmax = h), position = pos, size = 2) +
  geom_linerange(aes(xmin = ll, xmax = hh), position = pos) +
  geom_point(position = pos, color = "black")
```

Second model:

```{r posterors-2nd-model}
mcmc_intervals(
  posterior2,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
  ) +
  theme_classic()
```

@@

```{r effect-sizes}
#emmeans(m_1, revpairwise ~ target_type + sand_ratio | exposition |
#survey_year_fct, type = "response"))
```

>>>>

```{r load-models, collapse = TRUE, message = FALSE}
load(file = here("outputs", "models", "model_recovery_2.Rdata"))
load(file = here("outputs", "models", "model_recovery_full.Rdata"))
load(file = here("outputs", "models", "model_recovery_2_prior.Rdata"))
# BARG 5.A/B/C
load(file = here("outputs", "models", "model_recovery_2_flat.Rdata"))
m_1 <- m2
m_2 <- m_full
m_prior <- m2_prior
m_flat <- m2_flat
```

@@

```{r histograms}
p1 <- ppc_stat(y, yrep1, binwidth = 0.001)
p2 <- ppc_stat(y, yrep2, binwidth = 0.001)
p1 / p2
ppc_stat_grouped(y, yrep1, group = sites$site, binwidth = 0.001)
ppc_stat_grouped(y, yrep2, group = sites$site, binwidth = 0.001)
p1 <- ppc_stat_grouped(y, yrep1, group = sites$exposition, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$exposition, binwidth = 0.001)
p1 / p2
ppc_stat_grouped(y, yrep1, group = sites$survey_year_fct, binwidth = 0.001)
ppc_stat_grouped(y, yrep2, group = sites$survey_year_fct, binwidth = 0.001)
p1 <- ppc_stat_grouped(y, yrep1, group = sites$target_type, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$target_type, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$seed_density, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$seed_density, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$sand_ratio, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$sand_ratio, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$substrate_depth, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$substrate_depth, binwidth = 0.001)
p1 / p2
```

@@

```{r loo}
loo1
loo2
plot(loo1)
plot(loo2)
```

@@

```{r loo-pit}
p1 <- bayesplot::ppc_loo_pit_overlay(y, yrep1, lw = weights(loo1$psis_object))
p2 <- bayesplot::ppc_loo_pit_overlay(y, yrep2, lw = weights(loo2$psis_object))
p1 / p2

@@

```{r posteriors, fig.height = 7}
combined <- rbind(
  bayesplot::mcmc_intervals_data(posterior1, prob = 0.66, prob_outer = 0.95),
  bayesplot::mcmc_intervals_data(posterior2, prob = 0.66, prob_outer = 0.95),
  bayesplot::mcmc_intervals_data(posterior_prior, prob = 0.66, prob_outer = 0.95),
  bayesplot::mcmc_intervals_data(posterior_flat, prob = 0.66, prob_outer = 0.95)
  )
combined$model <- rep(
  c("m_1", "m_2", "m_flat", "m_prior"),
  each = posterior1$"1" %>% as_tibble() %>% ncol()
  )
pos <- position_nudge(
  y = if_else(
    combined$model == "m_2", -.2, if_else(
      combined$model == "m_prior", -.4, if_else(
        combined$model == "m_flat", -.6, 0
        )
      )
    )
  )
ggplot(data = combined, aes(x = m, y = forcats::fct_rev(factor(parameter)), color = model)) + 
  geom_vline(xintercept = 0, color = "grey") +
  geom_linerange(aes(xmin = l, xmax = h), position = pos, linewidth = 2) +
  geom_linerange(aes(xmin = ll, xmax = hh), position = pos) +
  geom_point(position = pos, color = "black") +
  coord_cartesian(xlim = c(-1, .6)) +
  bayesplot::theme_default() +
  ggtitle("Posterior distbributions (mean, CI66, CI95)")
```

@@

```{r effect-sizes}
emmeans(m_1, revpairwise ~ target_type + sand_ratio | exposition |
          survey_year_fct, type = "response")
```
